---
title: "Dominican Inflation Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ipc_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(databcrd)
library(ggplot2)
library(dplyr)
```

Inflation is undoubtedly one of the most important macroeconomic variables and the core 
of the monetary policy in the Dominican Republic. Its evolution affects the stability 
of real variables such as consumption and investment, attracting analysts' attention 
and being frequently consulted.

This article demonstrates how to quickly access inflation expectations in the 
Dominican Republic using the `{databcrd}` package, specifically the `get_ipc_data()` function.

## Published Inflation Breakdowns

1. General: National inflation without breakdowns
2. By groups: Breakdown by groups of goods and services
3. By regions: Inflation specific to the country's macro-regions
4. By components: Inflation of tradable and non-tradable goods
5. Core: Core inflation, excluding volatile items
6. Items: 

With the package, all inflation breakdowns can be accessed using the 
`get_ipc_data()` function. Simply specify the `desagregacion` argument. 
The possible values are: `"general"`, `"grupos"`, `"subyacente"`, `"regiones"`, `"tnt"`, 
and `"articulos"`.

## General Inflation

With `get_ipc_data("general")`, the updated general inflation data is downloaded, 
including the index (`ipc`), monthly variation (`ipc_vm`), year-over-year variation (`ipc_vi`), 
variation since December (`ipc_vd`), and the 12-month average variation (`ipc_p12`).

```{r showcase_get_ipc}
inflacion_general <- get_ipc_data("general")
inflacion_general
```

Let's generate graphs for each variable.

```{r plot_function}
# Function to plot inflation data
plot_ipc_data <- function(data, variable, title, start_year = 2018) {
  data |>
    filter(year == start_year) |> 
    ggplot(aes(x = fecha, y = {{ variable }})) +
    geom_line() +
    theme_minimal() +
    ggtitle(title) +
    theme(
      axis.title = element_blank(),
      plot.title.position = "plot"
    )
}
```

```{r plot_ipc, fig.width=6, fig.height=3, fig.align='center'}
plot_ipc_data(inflacion_general, ipc, "Consumer Price Index")
```

```{r plot_ipc_vm, fig.width=6, fig.height=3, fig.align='center'}
plot_ipc_data(inflacion_general, ipc_vm, "Monthly Variation of CPI") +
  scale_y_continuous(labels = \(x) scales::comma(x, accuracy = 0.1, suffix = "%"))
```

```{r plot_ipc_vis, fig.width=6, fig.height=3, fig.align='center'}
plot_ipc_data(inflacion_general, ipc_vi, "Year-over-Year Variation of CPI") +
  scale_y_continuous(labels = \(x) scales::comma(x, accuracy = 0.1, suffix = "%"))
```

## Now a _naive_ forecast of inflation

```{r forecasting}
library(tsibble)
library(fable)
library(feasts)

ts_ipc <- inflacion_general |>
  mutate(fecha = yearmonth(fecha)) |>
  select(-mes) |>
  as_tsibble(index = fecha)

models <- ts_ipc |>
  model(
    ets   = ETS(box_cox(ipc_vm, 0.3)),
    arima = ARIMA(ipc_vm),
    naive = SNAIVE(ipc_vm)
  )

models
```

```{r forecast_plots, fig.width=6, fig.height=3, fig.align='center'}
models |>
  forecast(h = "6 months") %>% 
  autoplot(filter(ts_ipc, year > 2018), level = NULL) +
  theme_minimal()
